---
title: "MEA_beatrate_peakhight"
author: "RvH"
date: "2-9-2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(drc)
library(tidyverse)
library(cowplot)
theme_set(theme_cowplot())
library(ggsci)
#theme_temp <- theme_cowplot()
#theme_temp <- theme_update(panel.grid.major = element_line(color = "gray90", size = 0.5))
#setwd("D:/OneDrive - LUMC/PhD/Writing/First/Figure 5 Fatty acids and extra/MEA/")
#setwd("Y:/OneDrive_LUMC/PhD/Writing/First/Figure 5 Fatty acids and extra/MEA/")
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}
number_axis <- list(theme(text = element_text(size=29.1), 
        axis.text.x = element_text(size = 22.8),
        axis.text.y = element_text(size = 22.8),
        title = element_text(size = 29.1)))
label_axis <-list(theme(text = element_text(size=29.1), 
        axis.text.x = element_text(size = 29.1),
        axis.text.y = element_text(size = 22.8),
        title = element_text(size = 29.1)))
```

##Import data
```{r}
load("Data/Amplitude_beatperiod_plate1.Rdata")
dat_all <- dat %>% mutate(exp = "Plate 1")
dat_mean_all <- dat_mean %>% mutate(exp = "Plate 1")
flec_all <- flec %>% mutate(exp = "Plate 1")
load("Data/Amplitude_beatperiod_plate3.Rdata")
dat_all <- bind_rows(dat_all, dat %>% mutate(exp = "Plate 3"))
dat_mean_all <- bind_rows(dat_mean_all, dat_mean %>% mutate(exp = "Plate 3"))
flec_all <- bind_rows(flec_all, flec %>% mutate(exp = "Plate 3"))

```

## Calculate mean statistics and plot data per well
```{r}

model1 <- drm(Amplitude_change ~ Concentration, fct = LL.4(names = c("Slope", "Lower", "Upper", "ED50")), data = flec_all %>% filter(Line == "Pt 1"))  
summary(model1)
plot(model1)

model2 <- drm(Amplitude_change ~ Concentration, fct = LL.4(names = c("Slope", "Lower", "Upper", "ED50")), data = flec_all %>% filter(Line == "Iso 1"))  
summary(model2)
plot(model2)

model3 <- drm(Amplitude_change ~ Concentration, fct = LL.4(names = c("Slope", "Lower", "Upper", "ED50")), data = flec_all %>% filter(Line == "Ctrl"))  
summary(model3)
plot(model3)

vert <- data.frame(Line = c("Pt 1", "Iso 1", "Ctrl"), EC50 = c(model1$fit$par[4], model2$fit$par[4], model3$fit$par[4]))
plot_flec <- flec_all %>% mutate(Concentration = ifelse(Concentration == 0, 0.01, Concentration))
ggplot(plot_flec, aes(x = Concentration, y = Amplitude_change, colour = Line))+
  stat_summary(fun = mean, geom = "point", shape = 1, size = 3)+
  geom_smooth(method = drm, method.args = list(fct= L.4()), se = FALSE)+
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1, size = 0.5)+
  scale_x_log10()+
  geom_vline(data = vert, aes(xintercept = EC50, colour = Line), linetype = "dashed")+
    scale_colour_npg()+
  number_axis+
  labs(y = "Amplitude decrease (%) ",
       x = "Concentration Flecainide (\u00b5M)")

save_plot("Graphs/Amplitude_flec.pdf", last_plot(), base_height = 6, base_asp = 1.2)
save_plot("Graphs/Amplitude_flec.png", last_plot(), base_height = 6, base_asp = 1.2)

ggplot(plot_flec %>% filter(file == "Baseline"), aes(x = Line, y = `Beat Period Mean (s)`,  fill = Line))+
    stat_summary(fun = mean, geom = "bar")+
  stat_summary(fun.data = mean_sdl, geom = "errorbar", size = 0.5, width = 0.1)+
  scale_y_continuous(expand = c(0,0))

ggplot(plot_flec %>% filter(file == "Baseline"), aes(x = Line, y = `Spike Slope Mean (V / s)`,  fill = Line))+
    stat_summary(fun = mean, geom = "bar")+
  stat_summary(fun.data = mean_sdl, geom = "errorbar", size = 0.5, width = 0.1)+
  scale_y_continuous(expand = c(0,0))

ggplot(data = flec_all %>% mutate(Concentration = as.factor(Concentration)), aes(x = Concentration, y = Beat_rate_increase, fill = Line))+
   stat_summary(fun = mean, geom = "bar", position="dodge", stat="identity")+
  stat_summary(fun.data = mean_sdl, geom = "errorbar", size = 0.5, width = 0.1, position = position_dodge(0.8))+
  scale_y_continuous(expand = c(0,0))+  
  scale_fill_npg()

save(dat, flec, dat_mean, file = "Amplitude_beatperiod.Rdata")
```

## calculate mean statistics and plot data per well and electrode

dat_mean <- dat %>% group_by(Line, Well, Treatment, Electrode) %>% 
  mutate(baseline = `Spike Amplitude Mean (mV)`,
         Amplitude_change = `Spike Amplitude Mean (mV)`/`Spike Amplitude Mean (mV)`[file == "Baseline"] * 100,) %>% 
  select(1:3, Amplitude_change, baseline, everything()) %>% 
  group_by(file, Line, Well, Treatment) %>% summarise(across(where(is.double), ~mean(.x, na.rm=TRUE))) %>% 
  ungroup() 

flec <- dat_mean %>% filter(Treatment == "Flec") %>% ungroup() %>% 
  arrange(Line, file) %>% mutate(Concentration = c(rep(c(0,0.1,0.3,1,3,10,30), each = 3, times = 2), 
                                          rep(c(0,0.1,0.3,1,3,10,30), each = 2))) %>% select(1:3, Concentration, everything())
flec

model1 <- drm(Amplitude_change ~ Concentration, fct = LL.4(names = c("Slope", "Lower", "Upper", "ED50")), data = flec %>% filter(Line == "Pt 1"))  
summary(model1)
plot(model1)

model2 <- drm(Amplitude_change ~ Concentration, fct = LL.4(names = c("Slope", "Lower", "Upper", "ED50")), data = flec %>% filter(Line == "Iso 1"))  
summary(model2)
plot(model2)

model3 <- drm(Amplitude_change ~ Concentration, fct = LL.4(names = c("Slope", "Lower", "Upper", "ED50")), data = flec %>% filter(Line == "Ctrl"))  
summary(model3)
plot(model3)

DMSO <- dat_mean %>% filter(Treatment == "DMSO") %>% ungroup() %>% 
  arrange(Line, file) %>% mutate(Concentration = c(rep(c(0,0.0002,0.000667,0.002,0.00667,0.02,0.0667), each = 2), 
                                          rep(c(0,0.0002,0.000667,0.002,0.00667,0.02,0.0667), each = 2, times = 2))) %>% select(1:3, Concentration, everything())
DMSO

model <- drm(Amplitude_change ~ Concentration, fct = LL.4(names = c("Slope", "Lower", "Upper", "ED50")), data = DMSO %>% filter(Line == "Pt 1"))  
summary(model)
plot(model)

model <- drm(Amplitude_change ~ Concentration, fct = LL.4(names = c("Slope", "Lower", "Upper", "ED50")), data = DMSO %>% filter(Line == "Iso 1"))  
summary(model)
plot(model)

model <- drm(Amplitude_change ~ Concentration, fct = LL.4(names = c("Slope", "Lower", "Upper", "ED50")), data = DMSO %>% filter(Line == "Ctrl"))  
summary(model)
plot(model)


vert <- data.frame(Line = c("Pt 1", "Iso 1", "Ctrl"), EC50 = c(model1$fit$par[4], model2$fit$par[4], model3$fit$par[4]))

ggplot(flec, aes(x = Concentration, y = Amplitude_change, colour = Line))+
  stat_summary(fun = mean, geom = "point", shape = 1, size = 3)+
  geom_smooth(method = drm, method.args = list(fct= L.4()), se = FALSE)+
  scale_x_log10()+
  geom_vline(data = vert, aes(xintercept = EC50, colour = Line))


ggplot(dat_mean %>% filter(file == "Baseline"), aes(x = Line, y = `Beat Period Mean (s)`,  fill = Line))+
    stat_summary(fun = mean, geom = "bar")+
  stat_summary(fun.data = mean_sdl, geom = "errorbar", size = 0.5, width = 0.1)+
  scale_y_continuous(expand = c(0,0))

seq_plot <- function(y) {
  plot <- ggplot(dat_mean, aes_string(x = "Line", y = paste0("`",y, "`"), fill = "Line"))+
  stat_summary(fun = mean, geom= "bar")+
  geom_point()+
  stat_summary(fun.data = mean_se, geom = "errorbar")  
  plot(plot)
}

plot_vars <- names(dat_mean)[!str_detect(names(dat_mean), "sd")]

sapply(plot_vars, seq_plot)