---
title: "MEA_beatrate_peakhight"
author: "RvH"
date: "`r Sys.Date()`"
output: html_document
---

# First MEA test with Drug exposures
Goal of the experiment was to test several compounds which should elicit a strong effect on the FPD of the Cardioids.
For this purpose three drugs were chosen, Flecanaide, E-4031 and BayK, all in a concentration range from 0.001 - 3 uM.


```{r setup, include=FALSE}
#knitr::opts_knit$set(root.dir = "Y:/OneDrive_LUMC/PhD/Calcium/210830_MEA_first_analysis/Data")
#knitr::opts_knit$set(root.dir = "D:/OneDrive - LUMC/PhD/MEA/Plate 3/")
knitr::opts_chunk$set(echo = TRUE)
library(drc)
library(tidyverse)
library(cowplot)
theme_set(theme_cowplot())
library(ggsci)
#theme_temp <- theme_cowplot()
#theme_temp <- theme_update(panel.grid.major = element_line(color = "gray90", size = 0.5))
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}
setwd("C:/Users/ruben.vanhelden/OneDrive - HeartBeat.bio AG/MEA/230817_Drug_test/111-8253")
```

##Import data
```{r}
files <- list.files("Export1/", full.names = TRUE, pattern = ").csv")
if(!file.exists("beatrate_peakthight.rData")){

dat <- data.frame()
for(i in files){
df <- read_csv(i, skip = 154, n_max = 20) ##Select proper part of excelsheet
df <- df %>% pivot_longer(2:769, names_to = "Well", values_to = "blah") %>% 
  pivot_wider(names_from = `Measurement`, values_from = blah) %>% 
  filter(`Number Of Beats` > 0) %>% 
  mutate(file = substr(i, 9, nchar(i))) %>% 
  separate(Well,  into = c("Well", "Electrode"))
dat <- rbind(dat,df)
print(paste("merging dat and", i))
}
dat <- dat %>% mutate(Settings =  case_when( ##extract name from file and create groups
          str_detect(file, "(001)") ~ "Baseline", 
          str_detect(file, "addition_1") ~ "Dose 1", 
          str_detect(file, "80_min") ~ "Dose 2_80 min", 
          TRUE ~ "Dose 2"
          )) %>% select(file, Settings, everything())
save(dat, file = "beatrate_peakthight.rData") 
} else {
  load("beatrate_peakthight.rData")
}

Plate_layout <- read_csv("Plate_layout.csv")
dat <- right_join(dat, Plate_layout, by = c("Well")) %>% ##merge with plate layout
  mutate(Concentration = case_when(
    str_detect(Settings, "Dose 1") ~ Concentration.1, 
    str_detect(Settings, "Dose 2") ~ Concentration.2,
    TRUE ~ 0))
dat <- dat %>% filter(Settings == "Baseline") %>% select(Well, Electrode) %>% left_join(dat) %>% ##select only baseline active electrodes
  filter(Settings == "Dose 2") %>% select(Well, Electrode) %>% left_join(dat) ##select all wells which are still active at dose 2

dat <- dat %>% select(file, Settings, Well, Drug, Concentration, everything())

glimpse(dat)
dat
```

## Filter correct data for experiments
Verify that each well has active electrodes that do not diverge too much from eachother
```{r}

ggplot(dat %>% filter(Drug == "DMSO"), aes(x = Concentration, y = `Number Of Beats`, colour = Electrode))+
  geom_line()+
  facet_wrap(~Well)

ggplot(dat %>% filter(Drug == "Flec"), aes(x = Concentration, y = `Number Of Beats`, colour = Electrode))+
  geom_line()+
  facet_wrap(~Well)

#C10 misses electrode 22 at concentration 0.1 and electrode at concentration 23 long dose. Does not seem to give deviating data at other points, will be included.
#C9 Electrode 23 and 31 give deviating number of beats at dose 1 and 2. Will be excluded from those time points

ggplot(dat %>% filter(Drug == "E-4031"), aes(x = Concentration, y = `Number Of Beats`, colour = Electrode))+
  geom_line()+
  facet_wrap(~Well)

#E10 has deviating data in electrode 12 and 22 at Dose 2 and dose 2 80 min, will be excluded from those time points
#F10 should reinspect the traces, only 2 electrodes give data which deviates from eachother. For now selecting electrode 13 which is closest to baseline mean of other wells.
#E7 electode 33 deviates from all other at all timepoints, should be removed.
#E6 electrode 12 and 31 deviate at dose 2 and dose 2 80 minutes, will be removed from these timepoints
#E9 electrode 21 has issues at baseline, remove from all timepoints
#E12 Electrode 31 deviates in Dose 2 and dose 2 80 minutes

ggplot(dat %>% filter(Drug == "BayK"), aes(x = Concentration, y = `Number Of Beats`, colour = Electrode))+
  geom_line()+
  facet_wrap(~Well)

#G5 electrode 11 deviates at baseline, will be removed from all timepoints
#H8 multiple electrodes deviate at all time points, well will be completely removed.
#H10 electrode 23 deviates at baseline, remove completely 

dat <- dat %>% 
  filter(!(Well == "C9" & Electrode %in% c(23,31) & Settings %in% c("Dose 1", "Dose 2"))) %>% 
  filter(!(Well == "E10" & Electrode %in% c(12, 22) & Settings %in% c("Dose 2", "Dose 2_80 min"))) %>%
  filter(!(Well == "F10" & Electrode %in% c(12))) %>%
  filter(!(Well == "E7" & Electrode %in% c(33))) %>%
  filter(!(Well == "E6" & Electrode %in% c(12,31) & Settings %in% c("Dose 2", "Dose 2_80 min"))) %>%
  filter(!(Well == "E9" & Electrode %in% c(21))) %>%
  filter(!(Well == "E12" & Electrode %in% c(31) & Settings %in% c("Dose 2", "Dose 2_80 min"))) %>%
  filter(!(Well == "G5" & Electrode %in% c(11))) %>%
  filter(!(Well == "H8")) %>%
    filter(!(Well == "H10" & Electrode %in% c(23))) #%>%
  
## all wells have SD below 5
dat %>% group_by(file, Drug, Concentration, Well) %>% ##Shows baseline electrodes  still with issues 
  summarise(Beats = mean(`Number Of Beats`), Beats_median = median(`Number Of Beats`), Beats_sd = sd(`Number Of Beats`), m_Beats = getmode(`Number Of Beats`)) %>% 
  arrange(desc(Beats_sd)) #%>% filter(file == "base")

```


## Look at difference in dosing time interval (10 min vs 80 min) and DMSO effect

```{r}
df <- dat 
df$Conc <- factor(df$Concentration, levels = sort(unique(df$Concentration)))

bplot <- function(y, drg){
plot <- ggplot(df1 %>% filter(Drug == drg), aes(x = Settings, y = .data[[y]], fill = Conc))+
  geom_boxplot()+
  labs(title = drg)
plot(plot)
}

df1 <- df %>% 
  group_by(Well, Drug) %>% 
    mutate(baseline = `Spike Amplitude Mean (mV)`,
         Amplitude_change = `Spike Amplitude Mean (mV)`/mean(`Spike Amplitude Mean (mV)`[Concentration == 0]) * 100,
         Beat_rate_increase = `Number Of Beats`/mean(`Number Of Beats`[Concentration == 0])*100,
         FPD_change = `FPD Mean (ms)`/mean(`FPD Mean (ms)`[Concentration == 0])*100, 
         Slope_change = `Spike Slope Mean (V / s)`/mean(`Spike Slope Mean (V / s)`[Concentration == 0])*100) %>% 
  select(1:3, Amplitude_change, Beat_rate_increase, FPD_change, Slope_change, baseline, everything())


nams <- names(df1)[4:7]
sapply(nams, bplot, "DMSO")
sapply(nams, bplot, "BayK")
sapply(nams, bplot, "Flec")
sapply(nams, bplot, "E-4031")

  
pairwise.t.test(df1$Beat_rate_increase[df1$Drug == "DMSO"], df1$Concentration[df1$Drug == "DMSO"], p.adj = "none")
pairwise.t.test(df1$Beat_rate_increase[df1$Drug == "Flec"], df1$Concentration[df1$Drug == "Flec"], p.adj = "none")
```


## Calculate mean statistics, uncorrected for DMSO
```{r}
dat_mean <- dat %>% group_by(file, Well, Drug, Settings) %>% summarise(across(where(is.double), ~mean(.x, na.rm=TRUE))) %>% 
  ungroup() %>% group_by(Well, Drug) %>% 
  mutate(baseline = `Spike Amplitude Mean (mV)`,
         Amplitude_change = `Spike Amplitude Mean (mV)`/`Spike Amplitude Mean (mV)`[Concentration == 0] * 100,
         Beat_rate_increase = `Number Of Beats`/`Number Of Beats`[Concentration == 0]*100,
         FPD_change = `FPD Mean (ms)`/`FPD Mean (ms)`[Concentration == 0]*100, 
         Slope_change = `Spike Slope Mean (V / s)`/`Spike Slope Mean (V / s)`[Concentration == 0]*100) %>% 
  select(1:3, Amplitude_change, Beat_rate_increase, FPD_change, Slope_change, baseline, everything())


dat_mean
glimpse(dat_mean)

Slope <- dat_mean %>% filter(! Settings == "Dose 2_80 min") %>% 
  drc::drm(data=., Slope_change ~ Concentration, fct = drc::LL.4(names = c("Slope", "Lower", "Upper", "ED50")), curveid = Drug) 
Amplitude <- dat_mean%>% filter(! Settings == "Dose 2_80 min",! Drug %in% c("DMSO", "E-4031")) %>% 
  drc::drm(data=., Amplitude_change ~ Concentration, fct = drc::LL.4(names = c("Slope", "Lower", "Upper", "ED50")), curveid = Drug) 
BR <- dat_mean %>% filter(! Settings == "Dose 2_80 min",! Drug == "DMSO") %>% 
  drc::drm(data=., Beat_rate_increase ~ Concentration, fct = drc::LL.4(names = c("Slope", "Lower", "Upper", "ED50")), curveid = Drug) 
FPD <- dat_mean%>% filter(! Settings == "Dose 2_80 min",! Drug == "DMSO") %>% 
  drc::drm(data=., FPD_change ~ Concentration, fct = drc::LL.4(names = c("Slope", "Lower", "Upper", "ED50")), curveid = Drug) 
  
plot(Slope)
plot(Amplitude)
plot(BR)
plot(FPD)

dat_mean <- dat_mean %>% mutate(Conc = ifelse(Concentration == 0, 0.0001, Concentration))
dat <- dat_mean

splot <- function(dat, drg, y){
plot <- ggplot(dat %>% filter(Drug %in% drg), aes(x = Conc, y = .data[[y]], colour = Drug))+
  stat_summary(fun = mean, geom = "point", shape = 1, size = 3)+
  geom_point()+
  geom_smooth(method = drm, method.args = list(fct= L.4()), se = FALSE)+
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1, size = 0.5)+
  scale_x_log10()+
#plot
#  geom_vline(data = vert, aes(xintercept = EC50, colour = Line), linetype = "dashed")+
#    scale_colour_npg()+
  labs(#y = "Amplitude decrease (%) ",
       x = "Concentration Flecanaide (uM)")
plot(plot)
}

splot(dat_mean, c("E-4031", "BayK", "Flec"), "Amplitude_change")
splot(dat_mean, c("E-4031", "BayK", "Flec"), "Slope_change")
splot(dat_mean, c("E-4031", "BayK", "Flec"), "Beat_rate_increase")
splot(dat_mean, c("DMSO"), "Beat_rate_increase")
```






## calculate mean statistics and plot data per well and electrode

dat_mean <- dat %>% group_by(Line, Well, Treatment, Electrode) %>% 
  mutate(baseline = `Spike Amplitude Mean (mV)`,
         Amplitude_change = `Spike Amplitude Mean (mV)`/`Spike Amplitude Mean (mV)`[file == "Baseline"] * 100,) %>% 
  select(1:3, Amplitude_change, baseline, everything()) %>% 
  group_by(file, Line, Well, Treatment) %>% summarise(across(where(is.double), ~mean(.x, na.rm=TRUE))) %>% 
  ungroup() 

flec <- dat_mean %>% filter(Treatment == "Flec") %>% ungroup() %>% 
  arrange(Line, file) %>% mutate(Concentration = c(rep(c(0,0.1,0.3,1,3,10,30), each = 3, times = 2), 
                                          rep(c(0,0.1,0.3,1,3,10,30), each = 2))) %>% select(1:3, Concentration, everything())
flec

model1 <- drm(Amplitude_change ~ Concentration, fct = LL.4(names = c("Slope", "Lower", "Upper", "ED50")), data = flec %>% filter(Line == "Pt 1"))  
summary(model1)
plot(model1)

model2 <- drm(Amplitude_change ~ Concentration, fct = LL.4(names = c("Slope", "Lower", "Upper", "ED50")), data = flec %>% filter(Line == "Iso 1"))  
summary(model2)
plot(model2)

model3 <- drm(Amplitude_change ~ Concentration, fct = LL.4(names = c("Slope", "Lower", "Upper", "ED50")), data = flec %>% filter(Line == "Ctrl"))  
summary(model3)
plot(model3)

DMSO <- dat_mean %>% filter(Treatment == "DMSO") %>% ungroup() %>% 
  arrange(Line, file) %>% mutate(Concentration = c(rep(c(0,0.0002,0.000667,0.002,0.00667,0.02,0.0667), each = 2), 
                                          rep(c(0,0.0002,0.000667,0.002,0.00667,0.02,0.0667), each = 2, times = 2))) %>% select(1:3, Concentration, everything())
DMSO

model <- drm(Amplitude_change ~ Concentration, fct = LL.4(names = c("Slope", "Lower", "Upper", "ED50")), data = DMSO %>% filter(Line == "Pt 1"))  
summary(model)
plot(model)

model <- drm(Amplitude_change ~ Concentration, fct = LL.4(names = c("Slope", "Lower", "Upper", "ED50")), data = DMSO %>% filter(Line == "Iso 1"))  
summary(model)
plot(model)

model <- drm(Amplitude_change ~ Concentration, fct = LL.4(names = c("Slope", "Lower", "Upper", "ED50")), data = DMSO %>% filter(Line == "Ctrl"))  
summary(model)
plot(model)


vert <- data.frame(Line = c("Pt 1", "Iso 1", "Ctrl"), EC50 = c(model1$fit$par[4], model2$fit$par[4], model3$fit$par[4]))

ggplot(flec, aes(x = Concentration, y = Amplitude_change, colour = Line))+
  stat_summary(fun = mean, geom = "point", shape = 1, size = 3)+
  geom_smooth(method = drm, method.args = list(fct= L.4()), se = FALSE)+
  scale_x_log10()+
  geom_vline(data = vert, aes(xintercept = EC50, colour = Line))


ggplot(dat_mean %>% filter(file == "Baseline"), aes(x = Line, y = `Beat Period Mean (s)`,  fill = Line))+
    stat_summary(fun = mean, geom = "bar")+
  stat_summary(fun.data = mean_sdl, geom = "errorbar", size = 0.5, width = 0.1)+
  scale_y_continuous(expand = c(0,0))

seq_plot <- function(y) {
  plot <- ggplot(dat_mean, aes_string(x = "Line", y = paste0("`",y, "`"), fill = "Line"))+
  stat_summary(fun = mean, geom= "bar")+
  geom_point()+
  stat_summary(fun.data = mean_se, geom = "errorbar")  
  plot(plot)
}

plot_vars <- names(dat_mean)[!str_detect(names(dat_mean), "sd")]

sapply(plot_vars, seq_plot)