---
title: "MEA_analysis"
author: "RvH"
date: "1-9-2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_knit$set(root.dir = "D:/OneDrive - LUMC/PhD/Calcium/210830_MEA_first_analysis/Data")0
#knitr::opts_knit$set(root.dir = "X:/OneDrive_LUMC/PhD/MEA/Plate 3/")
#library(drc)
library(tidyverse)
library(cowplot)
theme_set(theme_cowplot(font_size = 18))
#library(ggsci)
#library(ggrepel)
std <- function(x, na.rm = FALSE) sd(x, na.rm = na.rm)/sqrt(length(x))
#theme_temp <- theme_cowplot()
#theme_temp <- theme_update(panel.grid.major = element_line(color = "gray90", size = 0.5))
#setwd("D:/OneDrive - LUMC/PhD/Writing/First/Figure 5 Fatty acids and extra/MEA")

number_axis <- list(theme(text = element_text(size=29.1), 
        axis.text.x = element_text(size = 22.8),
        axis.text.y = element_text(size = 22.8),
        title = element_text(size = 29.1)))
label_axis <-list(theme(text = element_text(size=29.1), 
        axis.text.x = element_text(size = 29.1),
        axis.text.y = element_text(size = 22.8),
        title = element_text(size = 29.1)))
```

## build dataframe
```{r}
df <- tibble()
i <- 0 
while (i < 100){
df <- bind_rows(df, 
                read_csv("Batch_230818_RvH23D002_MEA_drug_test_D30(001)(000)_FP_CardiacMetrics.csv",
                         col_names = c("Measurement", sprintf(paste0(rep(LETTERS[1:8],each=12),"%01d"),1:12)),
                         skip = 5+i,
                         n_max = 14) %>% 
                  mutate(file = 
                           read_csv("Batch_230818_RvH23D002_MEA_drug_test_D30(001)(000)_FP_CardiacMetrics.csv",
                                    skip = i,
                                    n_max = 1,
                                    col_names = F) %>% pull(X2),
                         Settings =  case_when( ##extract name from file and create groups
                           str_detect(file, "(001)") ~ "Baseline", 
                           str_detect(file, "addition_1") ~ "Dose 1", 
                           str_detect(file, "80_min") ~ "Dose 2_80 min", 
                           TRUE ~ "Dose 2")) %>% 
                  select(file, Settings, everything()))
i = i + 25
}
Plate_layout <- read_csv("Plate_layout.csv")
df <- df %>% 
  pivot_longer(4:99, names_to = "Well", values_to = "x") %>% 
  pivot_wider(names_from = Measurement, values_from = x) %>% 
  right_join(Plate_layout, by = c("Well")) %>% ##merge with plate layout
  mutate(Concentration = case_when(
    str_detect(Settings, "Dose 1") ~ Concentration.1, 
    str_detect(Settings, "Dose 2") ~ Concentration.2,
    TRUE ~ 0)) %>% 
  select(file, Well, Settings, Concentration, everything()) %>% 
   relocate(file, .after = last_col())



load("Data/FPD_plate1.Rdata") ##already corrected for arrhythmia
df <- dat %>% mutate(exp = "plate1")
#load("Data/Analyzed_FPD_plate2.Rdata")
#df <- bind_rows(df, dat %>% mutate(exp = "plate2"))
load("Data/Analyzed_FPD_plate3.Rdata")
df <- bind_rows(df, dat %>% mutate(exp = "plate3")) %>% 
  unite("Measurement", 3:4, sep = "")

#Filter out all arrhythmic measures
arr <- read_csv("Data/arrhythmia_stats.csv") %>% filter(!is.na(`Treatment/ID`), !`Arrhythmia Flag` == 0) %>% 
  rename(exp = Plate)## Plate 2 corrected for naming mistake

df <- anti_join(df, arr, by = c("Display Name", "exp", "Measurement"))
```

## Models for E4031

```{r}
E4031 <- df %>% filter(Treatment == "E4031")

model1 <- drm(FPD_increase ~ Concentration, fct = LL.4(names = c("Slope", "Lower", "Upper", "ED50")), data = E4031 %>% filter(Line == "Pt 1"))  
summary(model1)
plot(model1)

model2 <- drm(FPD_increase ~ Concentration, fct = LL.4(names = c("Slope", "Lower", "Upper", "ED50")), data = E4031 %>% filter(Line == "Iso 1"))  
summary(model2)
plot(model2)

model3 <- drm(FPD_increase ~ Concentration, fct = LL.4(names = c("Slope", "Lower", "Upper", "ED50")), data = E4031 %>% filter(Line == "Ctrl"))  
summary(model3)
plot(model3)

DMSO <- df %>% filter(Treatment == "DMSO")

model <- drm(FPD_increase ~ Concentration, fct = LL.4(names = c("Slope", "Lower", "Upper", "ED50")), data = DMSO %>% filter(Line == "Pt 1"))  
summary(model)
plot(model)

model <- drm(FPD_increase ~ Concentration, fct = LL.4(names = c("Slope", "Lower", "Upper", "ED50")), data = DMSO %>% filter(Line == "Iso 1"))  
summary(model)
plot(model)

model <- drm(FPD_increase ~ Concentration, fct = LL.4(names = c("Slope", "Lower", "Upper", "ED50")), data = DMSO %>% filter(Line == "Ctrl"))  
summary(model)
plot(model)

vert <- data.frame(Line = c("Pt 1", "Iso 1", "Ctrl"), EC50 = c(model1$fit$par[4], model2$fit$par[4], model3$fit$par[4]))
plot_dat <- df %>% filter(Treatment == "E4031") %>% mutate(Concentration = ifelse(Concentration == 0, 3e-04, Concentration))
ggplot(plot_dat, aes(x = Concentration, y = FPD_increase, colour = Line, fill = Line))+
  stat_summary(fun = mean, geom = "point", shape = 16, size = 3)+
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1, size = 0.5)+
  geom_smooth(method = drm, method.args = list(fct= L.4()), se = FALSE)+
  scale_x_log10()+
  geom_vline(data = vert, aes(xintercept = EC50, colour = Line), linetype = "dashed")+
  scale_colour_npg()+
  labs(y = "FPD increase (%)",
       x = "Concentration E4031 (\u00b5M)")+
  number_axis
   # geom_text_repel(data = vert, aes(x = EC50, y = 90, label = round(EC50,3)),
    #                colour = "black", size = 3.5) 


#save_plot("Graphs/FPD_E4031.pdf", last_plot(), base_height = 6, base_asp = 1.2)
df_mean <- df %>% group_by(Line) %>% filter(Concentration == 0) %>% 
  summarise(across(where(is.double), list(mean = mean, std = std, sd = sd)))

ggplot(df_mean, aes(x = Line, y = `FPD (ms)_mean`, fill = Line))+
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = `FPD (ms)_mean` - `FPD (ms)_std`,
                ymax = `FPD (ms)_mean` + `FPD (ms)_std`), width = .2,
                position=position_dodge(.9)) + 
  scale_y_continuous(expand = c(0,0))+  
  labs(y = "FPD (ms)")+
  scale_fill_npg(guide = "none")+
  label_axis 
   
#save_plot("Graphs/FPD_lines.pdf", last_plot(), base_height = 6, base_asp = 1)
save_plot("Graphs/FPD_lines.png", last_plot(), base_height = 6, base_asp = 1)
test <- df %>% filter(`Display Name` == "B", Line %in% c("Pt 1", "Iso 1", "Ctrl"))
test$Line <- as.factor(test$Line)
pairwise.t.test(test$`FPD (ms)`, test$Line, p.adjust.method = "none")

#t.test(`FPD (ms)`~ Line, data = test)

test2 <- df %>% filter(Line %in% c("Pt 1", "Iso 1"), Treatment == "E4031", Concentration > 0.3) %>% 
  unite("Lines", c("Line", "Concentration"), remove = FALSE)

wilcox.test(FPD_increase ~ Line, data = test2 %>% filter(Line %in% c("Pt 1", "Iso 1"), Treatment == "E4031", Concentration == 3), alternative = "less")

```


#### test
```{r}

flec <- df %>% filter(! `Display Name` == "C8", Treatment == "Flec") %>% 
  mutate(Concentration = ifelse(Concentration == 0, 0.01, Concentration))

ggplot(flec, aes(x = Concentration, y = FPD_increase, colour = Line))+
  stat_summary(fun = mean, geom = "point", shape = 16, size = 3)+
  #geom_smooth(method = drm, method.args = list(fct= L.4()), se = FALSE, linetype = "dashed")+
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1, size = 0.5)+
  stat_summary(fun = mean, geom = "line")+
 # scale_x_log10()+
  scale_colour_npg()+
  labs(y = "FPD change (%)",
       x = "Concentration Flecainide (\u00b5M)")+
  number_axis+
  scale_x_continuous(trans = "log10", labels = scales::label_comma(accuracy = 0.01, trim = FALSE))

#save_plot("graphs/Flec_FPD.pdf", last_plot(), base_asp = 1.2, base_height = 6)

save(df, E4031, flec, file = "Analyzed_FPD.Rdata")
```

arr <- dat %>% group_by(Line, Treatment, Concentration) %>% filter(Treatment == "E4031", !`Display Name` == "C8") %>% 
  summarise(number = n()) %>% pivot_wider(names_from = Concentration, values_from = number) %>% 
  mutate(baseline = as.double(`0`)) %>% mutate(across(where(is.integer), ~.x / baseline*100)) %>% select(-baseline) %>% 
  pivot_longer(3:10, names_to = "Concentration", values_to = "Percentage arrhythmia") 
arr[is.na(arr)] <- 0
ggplot(dat = arr, aes(x = Concentration, y = 100-`Percentage arrhythmia`, fill = Line))+
  geom_bar(position="dodge", stat="identity")+
  scale_fill_npg()+
  scale_y_continuous(expand = c(0,0), limit = c(0,100))+
  labs(y = "Percentage E4031 induced arrhythmia",
       x = "Concentration E4031 (\u00b5M)")

arr_flec <- dat %>% group_by(Line, Treatment, Concentration) %>% filter(Treatment == "Flec", !`Display Name` == "C8") %>% 
  summarise(number = n()) %>% pivot_wider(names_from = Concentration, values_from = number) %>% 
  mutate(baseline = as.double(`0`)) %>% mutate(across(where(is.integer), ~.x / baseline*100)) %>% select(-baseline) %>% 
  pivot_longer(3:10, names_to = "Concentration", values_to = "Percentage arrhythmia")
  



ggplot(dat = arr_flec, aes(x = Concentration, y = 100-`Percentage arrhythmia`, fill = Line))+
  geom_bar(position="dodge", stat="identity")+
  scale_fill_npg()+
  scale_y_continuous(expand = c(0,0), limit = c(0,100))+
  labs(y = "Percentage Flecainide induced arrhythmia",
       x = "Concentration Flecainide (\u00b5M)")


ggplot(data = dat %>% filter(`Display Name` == "B"), aes(x = Line, y = `Conduction Velocity (mm/ms)`, fill = Line))+
   stat_summary(fun = mean, geom = "bar")+
  stat_summary(fun.data = mean_sdl, geom = "errorbar", size = 0.5, width = 0.1)+
  scale_y_continuous(expand = c(0,0))+  
  scale_fill_npg(guide = FALSE)
  


