---
title: "MEA_analysis"
author: "RvH"
date: "1-9-2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_knit$set(root.dir = "D:/OneDrive - LUMC/PhD/Calcium/210830_MEA_first_analysis/Data")0
#knitr::opts_knit$set(root.dir = "X:/OneDrive_LUMC/PhD/MEA/Plate 3/")
library(drc)
library(tidyverse)
library(cowplot)
theme_set(theme_cowplot(font_size = 18))
#library(ggsci)
#library(ggrepel)
std <- function(x, na.rm = FALSE) sd(x, na.rm = na.rm)/sqrt(length(x))
#theme_temp <- theme_cowplot()
#theme_temp <- theme_update(panel.grid.major = element_line(color = "gray90", size = 0.5))
#setwd("D:/OneDrive - LUMC/PhD/Writing/First/Figure 5 Fatty acids and extra/MEA")

number_axis <- list(theme(text = element_text(size=29.1), 
        axis.text.x = element_text(size = 22.8),
        axis.text.y = element_text(size = 22.8),
        title = element_text(size = 29.1)))
label_axis <-list(theme(text = element_text(size=29.1), 
        axis.text.x = element_text(size = 29.1),
        axis.text.y = element_text(size = 22.8),
        title = element_text(size = 29.1)))
```

## build dataframe
```{r}
if(!file.exists("FPD_data.Rdata")) {
df <- tibble()
i <- 0 
while (i < 100){
df <- bind_rows(df, 
                read_csv("Batch_230818_RvH23D002_MEA_drug_test_D30(001)(000)_FP_CardiacMetrics.csv",
                         col_names = c("Measurement", sprintf(paste0(rep(LETTERS[1:8],each=12),"%01d"),1:12)),
                         skip = 5+i,
                         n_max = 14) %>% 
                  mutate(file = 
                           read_csv("Batch_230818_RvH23D002_MEA_drug_test_D30(001)(000)_FP_CardiacMetrics.csv",
                                    skip = i,
                                    n_max = 1,
                                    col_names = F) %>% pull(X2),
                         Settings =  case_when( ##extract name from file and create groups
                           str_detect(file, "(001)") ~ "Baseline", 
                           str_detect(file, "addition_1") ~ "Dose 1", 
                           str_detect(file, "80_min") ~ "Dose 2_80 min", 
                           TRUE ~ "Dose 2")) %>% 
                  select(file, Settings, everything()))
i = i + 25
}
Plate_layout <- read_csv("Plate_layout.csv")
df <- df %>% 
  pivot_longer(4:99, names_to = "Well", values_to = "x") %>% 
  pivot_wider(names_from = Measurement, values_from = x) %>% 
  right_join(Plate_layout, by = c("Well")) %>% ##merge with plate layout
  mutate(Concentration = case_when(
    str_detect(Settings, "Dose 1") ~ Concentration.1, 
    str_detect(Settings, "Dose 2") ~ Concentration.2,
    TRUE ~ 0)) %>% 
  select(file, Well, Settings, Concentration, everything()) %>% 
   relocate(file, .after = last_col()) %>% 
  filter(!is.na(`FPD (ms)`))

  save(df, file = "FPD_data.Rdata")
} else {
  load("FPD_data.Rdata")
}
```


## plot raw data 
plot the data uncorrected for baseline or DMSO
excluding data of the second dose after 80 minutes
```{r}
df <- df %>% select(1:2, Drug, everything()) %>% 
  mutate(ajd.conc = ifelse(Settings == "Dose 2_80 min", Concentration + 1, 
                           ifelse( Settings == "Baseline", 0.0001, Concentration)),
         BPM = 60/`Beat Period (s)`)


rplot <- function(y, d){
  plot <- ggplot(d, aes(x = ajd.conc, y = .data[[y]], colour = Drug))+
    stat_summary(fun = mean, geom = "point", shape = 1, size = 3)+
#    geom_point()+
    stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1, size = 0.5)+
    geom_smooth(method = drm, method.args = list(fct= L.4()), se = FALSE)+
    scale_x_log10()
  plot(plot)
}
nams <- names(df)[c(10:12, 14,16,23)]
sapply(nams,rplot,df %>% filter(!Settings == "Dose 2_80 min"))
  

```

## Correct data for baseline and plot 
Here we see a potential change in beat rate for BayK and an increase in FPD for BayK, E4031 and Flec.
excluding data of the second dose after 80 minutes
```{r}
df_base <- df %>% 
  group_by(Well, Drug) %>% 
    mutate(baseline = `Spike Amplitude (mV)`,
         Amplitude_change = `Spike Amplitude (mV)`/mean(`Spike Amplitude (mV)`[Concentration == 0]),
         BPM_increase = BPM/mean(BPM[Concentration == 0])*100,
         FPD_change = `FPD (ms)`/mean(`FPD (ms)`[Concentration == 0])*100, 
         FPDc_change = `FPDc (Fridericia ms)`/mean(`FPDc (Fridericia ms)`[Concentration == 0])*100, 
         Slope_change = `Spike Slope (V/s)`/mean(`Spike Slope (V/s)`[Concentration == 0])*100,
         Velocity_change = `Conduction Velocity (mm/ms)`/mean(`Conduction Velocity (mm/ms)`[Concentration == 0])*100) %>% 
  select(1:4, Amplitude_change, BPM_increase, FPD_change, FPDc_change, Slope_change, Velocity_change, everything())

nams_base <- names(df_base)[5:10]
sapply(nams_base, rplot, df_base %>% filter(!Settings == "Dose 2_80 min"))
```

## Correct data for DMSO
Substracting the effect of DMSO on the statistics
```{r}
df_DMSO <- df_base %>% 
  group_by(Settings) %>% 
    mutate(baseline = `Spike Amplitude (mV)`,
         Amplitude_change = Amplitude_change - mean(Amplitude_change[Drug == "DMSO"]),
         BPM_increase = BPM_increase - mean(BPM_increase[Drug == "DMSO"]),
         FPD_change = FPD_change - mean(FPD_change[Drug == "DMSO"]), 
         FPDc_change = FPDc_change - mean(FPDc_change[Drug == "DMSO"]), 
         Slope_change = Slope_change - mean(Slope_change[Drug == "DMSO"]),
         Velocity_change = Velocity_change - mean(Velocity_change[Drug == "DMSO"])) %>% 
  select(1:4, Amplitude_change, BPM_increase, FPD_change, FPDc_change, Slope_change, Velocity_change, everything())
nams_DMSO <- names(df_DMSO)[5:10]
sapply(nams_DMSO, rplot, df_DMSO %>% filter(!Settings == "Dose 2_80 min"))

##Continue here with proper fold-change calculations change.
```

